Coupling finite and boundary element methods to solve the Poisson--Boltzmann equation for electrostatics in molecular solvation

### 2023-09-12 results.tex
Updated results

Folllowing the data set proposed by Fenley and co-workers,\cite{HarrisBoschitcshFenley2013} available in \url{http://www.sb.fsu.edu/~mfenley/convergence/downloads/convergence_pqr_sets.tar.gz}, and later used by Wei {\it et al.}\cite{nguyen2017accurate} to assess the accuracy of different Poisson-Boltzmann solvers, we tested our method by computing binding energies. 
This data set consists of three family of structures: DNA-drug complexes, different versions of the barnase-barstar complex, and RNA-peptide complexes.

According to the arginine results in Figure \ref{fig:arg_constant_energy}, a mesh density of 24 vertices per \AA$^2$ gives a solvation energy that is close to the true solution for a realistic geometry. 
In that case, the difference in solvation energy between calculations using finite element grids generated from a surface mesh of 8.6 and 24.5 vertices per \AA$^2$ (a 3$\times$ refinement) is less than 1\%.
Then, we used surface meshes with 24 vertices per \AA$^2$ and a quality parameter of 1.1 in TetGen to generate the meshes used to compute binding energies in Fenley's data base.

Several structure in the data base contained voids inside the molecular region where a water molecule fits. 
In a finite difference description, these voids are naturally considered as pockets of solvent, with the permittivity of water ($\epsilon=\epsilon_2$).
Regardless if this is an appropriate physical model, we also placed the permittivity of bulk water inside cavities for comparison purposes.
In the FEM formalism, we had to detect the elements contained by cavities and use a piece-wise constant varying description of the permittivity inside the molecular region, which can be done without any modification in the formulation. 
A BEM-BEM formulation can also account for cavities this way,\cite{AltmanBardhanWhiteTidor09,CooperBardhanBarba2014} but involves certain adjustments in the stiffness matrix.

Figures \ref{fig:dna_drug}, \ref{fig:barnase_barstar}, and \ref{fig:rna_peptide} show the binding energies of structures of DNA-drug complexes, barnase-barstar complexes, and RNA-peptide complexes, respectively, presented in Fenley's data set.\cite{HarrisBoschitcshFenley2013} The data for CPB\cite{HarrisBoschitcshFenley2013} and MIBPB, Delphi, APBS and PBSA where taken from the literature,\cite{nguyen2017accurate} for the finest mesh reported. 
From Figures \ref{fig:dna_drug}, \ref{fig:barnase_barstar}, and \ref{fig:rna_peptide} it is evident that our FEM-BEM approach gives results that are within the range of the other software.

@@

Even though this scheme is capable of calculating the electrostatic potential in medium-to-large proteins, the largest test case in Table \ref{table:large_variable} (1igt) used up 30\% of the available RAM memory. 
%(\textcolor{red}{check and rewrite \textcolor{blue}{we are using around 50\% of memory}}). 
If we were aiming at larger structures, such as full viruses,\cite{MartinezETal2019,wang2021high} we would require the use of optimized fast algorithms\cite{wang2021exafmm,kailasa2023pyexafmm} and
parellelization of the storage and solver, alongside the necessary improvements in the conditioning of the system.

>>>>

Following the data set proposed by Fenley and co-workers,\cite{HarrisBoschitcshFenley2013} available in \url{http://www.sb.fsu.edu/~mfenley/convergence/downloads/convergence_pqr_sets.tar.gz}, and later used by Wei {\it et al.}\cite{nguyen2017accurate} to assess the accuracy of different Poisson-Boltzmann solvers, we tested our method by computing binding energies. 
This data set consists of three families of structures: DNA-drug complexes, different versions of the barnase-barstar complex, and RNA-peptide complexes. Following the problem setting for these data sets, we use $\epsilon_1=1$, $\epsilon_2=80$ and an inverse of the Debye length of $\kappa=0.10265\,\text{\AA}^{-1}$.

According to the arginine results in Figure \ref{fig:arg_constant_energy}, a mesh density of 24 vertices per \AA$^2$ gives a solvation energy that is close to the true solution for a realistic geometry. 
In that case, the difference in solvation energy between calculations using finite element grids generated from a surface mesh of 8.6 and 24.5 vertices per \AA$^2$ (a 3$\times$ refinement) is less than 1\%.
Then, we used surface meshes with 24 vertices per \AA$^2$ and a quality parameter of 1.1 in TetGen to generate the meshes used to compute binding energies in Fenley's database.

Several structures in the database contained voids inside the molecular region where a water molecule fits. 
In a finite difference description, these voids are naturally considered as pockets of solvent, with the permittivity of water ($\epsilon=\epsilon_2$).
Regardless if this is an appropriate physical model, we also placed the permittivity of bulk water inside cavities for comparison purposes.
In the FEM formalism, we had to detect the elements contained by cavities and use a piece-wise constant varying description of the permittivity inside the molecular region, which can be done without any modification in the formulation. 
A BEM-BEM formulation can also account for cavities this way,\cite{AltmanBardhanWhiteTidor09,CooperBardhanBarba2014} but involves certain adjustments in the stiffness matrix.

Figures \ref{fig:dna_drug}, \ref{fig:barnase_barstar}, and \ref{fig:rna_peptide} show the binding energies of structures of DNA-drug complexes, barnase-barstar complexes, and RNA-peptide complexes, respectively, presented in Fenley's data set.\cite{HarrisBoschitcshFenley2013} The data for CPB\cite{HarrisBoschitcshFenley2013} and MIBPB, Delphi, APBS and PBSA were taken from the literature,\cite{nguyen2017accurate} for the finest mesh reported. 
From Figures \ref{fig:dna_drug}, \ref{fig:barnase_barstar}, and \ref{fig:rna_peptide} it is evident that our FEM-BEM approach gives results that are within the range of the other software.

@@

Even though this scheme is capable of calculating the electrostatic potential in medium-to-large proteins, the largest test case in Table \ref{table:large_variable} (1igt) used up 30\% of the available RAM memory. 
%(\textcolor{red}{check and rewrite \textcolor{blue}{we are using around 50\% of memory}}). 
If we were aiming at larger structures, such as full viruses,\cite{MartinezETal2019,wang2021high} we would require the use of optimised fast algorithms\cite{wang2021exafmm,kailasa2023pyexafmm} and
parellelization of the storage and solver, alongside the necessary improvements in the conditioning of the system.
